### 2022-05-09 frontmatter.tex
add working hepthesis template

%% Title
\titlepage[of Homerton College]{%
  A dissertation submitted to the University of Cambridge\\ for the degree of Doctor of Philosophy}

%% Abstract
\begin{abstract}%[\smaller \thetitle\\ \vspace*{1cm} \smaller {\theauthor}]
  %\thispagestyle{empty}
  \LHCb is a \bphysics detector experiment which will take data at
  the \SI{14}{\Hz} \LHC accelerator at \CERN from 2007 onward\dots
\end{abstract}

%% Declaration
\begin{declaration}
  This dissertation is the result of my own work, except where explicit
  reference is made to the work of others, and has not been submitted
  for another qualification to this or any other university. This
  dissertation does not exceed the word limit for the respective Degree
  Committee.
  \vspace*{1cm}
  \begin{flushright}
    Andy Buckley
  \end{flushright}
\end{declaration}

%% Acknowledgements
\begin{acknowledgements}
  Of the many people who deserve thanks, some are particularly prominent,
  such as my supervisor\dots
\end{acknowledgements}

%% Preface
\begin{preface}
  This thesis describes my research on various aspects of the \LHCb
  particle physics program, centred around the \LHCb detector and \LHC
  accelerator at \CERN in Geneva.

  \noindent
  For this example, I'll just mention \ChapterRef{chap:SomeStuff}
  and \ChapterRef{chap:MoreStuff}.
\end{preface}

%% ToC
\tableofcontents

%% Strictly optional!
\frontquote{%
  Writing in English is the most ingenious torture\\
  ever devised for sins committed in previous lives.}%
  {James Joyce}
%% I don't want a page number on the following blank page either.
\thispagestyle{empty}