### 2022-05-09 frontmatter.tex
try to reduce template bloat

%% Title
\titlepage[of Homerton College]{%
  A dissertation submitted to the University of Cambridge\\ for the degree of Doctor of Philosophy}

%% Abstract
\begin{abstract}%[\smaller \thetitle\\ \vspace*{1cm} \smaller {\theauthor}]
  %\thispagestyle{empty}
  \atlas\ is a detector experiment at \cern. $14~\eV{T}$.
\end{abstract}

%% Declaration
\begin{declaration}
  This dissertation is the result of my own work, except where explicit
  reference is made to the work of others, and has not been submitted
  for another qualification to this or any other university. This
  dissertation does not exceed the word limit for the respective Degree
  Committee.
  \vspace*{1cm}
  \begin{flushright}
    Your Name
  \end{flushright}
\end{declaration}

%% Acknowledgements
\begin{acknowledgements}
  Of the many people who deserve thanks, some are particularly prominent,
  such as my supervisor\dots
\end{acknowledgements}

%% Preface
\begin{preface}
  This thesis describes my research on various hoozywotsists.

  \noindent
  For this example, I'll just mention Chapter~\ref{chapter:intro}.
  and Chapter~\ref{chapter:intro}..
\end{preface}

%% ToC
\tableofcontents

\thispagestyle{empty}