### 2022-05-13 main/draft.tex
add junk tex file

As $\sigma \rightarrow 0$, this tends towards the the significance-transformed,
sign-flipped Poisson maximum likelihood ratio
\begin{equation}
S_\mathrm{Poisson}(n, \mu, \sigma)
= \mathrm{sgn}(n - \mu) \sqrt{2\left(
    n\log\left(\frac{n}{\mu}\right) - n + \mu
\right)}
~.
\end{equation}