### 2022-09-27 main/chapter_context.tex
add pmns, write a little on SM

\begin{singlespacing}
\chapter{Context: theory and experiment}
\label{chapter:context}
%
\begin{epigraphs}
\qitem{%
The idea that theorems follow from the postulates does not correspond to simple
observation.
If the Pythagorean theorem were found to not follow from the postulates, we
would again search for a way to alter the postulates until it was true.
Euclid's postulates came from the Pythagorean theorem, not the other way.}%
{Richard~Hamming,
\textit{The Unreasonable Effectiveness of Mathematics},
1980~\cite{hamming1980unreasonable}}
\qitem{%
if observing outer space gives us a view of the past, observing inner space
would surely give us a glimpse into the future - would be interesting if NASA
made a telescope for that%
}%
{Ken~M,
\textit{Comment: Yahoo! News},
2012~\cite{kenm2012inner}}
\end{epigraphs}
\end{singlespacing}

\noindent Theory is needed to understand experimental data,
and data are needed to improve theoretical models.
To clarify the experimental work of this thesis, this chapter introduces its
theoretical and experimental contexts;
Section~\ref{sec:context_sm_and_beyond} presents the current culmination of
theoretical particle physics in the Standard Model, its limitations, and its
possible future directions including supersymmetric ideas, and
Section~\ref{sec:context_atlas} then describes the \atlas\ detector, with whose
data we hope refine the Standard Model into its improved future forms.

\section{The Standard Model and beyond}
\label{sec:context_sm_and_beyond}
\noindent Our best current understanding of nature's fundamental objects in
their observable behaviours as particles is encoded in the Standard Model,
which uses solid mathematical foundations to accurately describe the physics
of particle interactions across a broad domain.
But that domain is bounded, so the Standard Model is not an ultimate theory of
everything, and work continues to expand that boundary on both theoretical and
experimental fronts.
The current Standard Model is described in Section~\ref{sec:context_sm},
along with its limitations and possible future directions.

Supersymmetry an idea that has been applied to build plausible alternatives
to the Standard Model which may explain future experimental discoveries.
Since the main work of this thesis is a search for effects of supersymmetric
phenomena, supersymmetry and the specific supersymmetric models that we study
are presented in Section~\ref{sec:context_susy}.

\subsection{The Standard Model}
\label{sec:context_sm}
The current Standard Model features fermions
(of four types in three generations),
which interact through forces mediated by vector bosons
and the scalar Higgs boson.
The strong force is mediated by
Massless gluons $g$ mediate the strong force,
the massless photon $\gamma$ mediates the electromagnetic force,
and very massive $W^+$, $W^-$, and $Z$ bosons mediate the weak force.
These fundamental particles of the Standard Model are summarized in
Table~\ref{tab:context_particles_sm}.

% using https://www.tablesgenerator.com/latex_tables
\begin{table}[tp]
\centering
\begin{tabular}{rrc}
\multirow{4}{*}{fermions} & \multirow{2}{*}{quarks}      & \makebox[5em]{\hfill $u$ \hfill $c$ \hfill $t$ \hfill}                    \\
                          &                              & \makebox[5em]{\hfill $d$ \hfill $s$ \hfill $b$ \hfill}                    \\[1ex]
                          & \multirow{2}{*}{leptions}    & \makebox[5em]{\hfill $e$ \hfill $\mu$ \hfill $\tau$ \hfill}               \\
                          &                              & \makebox[5em]{\hfill $\nu_e$ \hfill $\nu_\mu$ \hfill $\nu_\tau$ \hfill} \\[3ex]
\multirow{2}{*}{bosons}   & strong                       & \makebox[7em]{\hfill $g$ \hfill}                                          \\[1ex]
                          & electroweak & \makebox[7em]{\hfill $W^\pm\!\!$ \hfill $Z$\hfill $\gamma$ \hfill $h$ \hfill} \\
\end{tabular}
\caption[Particles of the Standard Model]{%
Particles of the Standard Model.
All fermions are spin-$1/2$, and all bosons are spin-$1$ except for the
spin-$0$ Higgs boson $h$.
Fermions come in left- and right-chiralities, although the existence of
right-chiral neutrinos is not strictly confirmed.%
}
\label{tab:context_particles_sm}
\end{table}

As a gauge field theory, the Standard Model has the symmetry group
$SU\!(3)_C \times SU\!(2)_L \times U\!(1)$,
meaning that its lagrangian is invariant under local operations of that group
on its fields in certain representations.

Gluons correspond to the eight generators of $SU\!(3)_C$, whose subscript $_C$
indicates that they operate on colour-charge triplets of quarks.

Weak and electromagnetic forces are unified in electroweak interactions with
four bosons $W^1$, $W^2$, $W^3$, and $B$ from the
$SU\!(2)_L$ and $U\!(1)_Y$ symmetries respectively.
These would be massless if the symmetry were perfectly respected, but it is
`spontaneously' broken with the Higgs mechanism~\cite{
higgs1964broken,
englert1964broken
}%
\footnote{%
Historical accuracy prefers the name `Brout–Englert–Higgs mechanism', but
Peter~Higgs has reportedly suggested
``ABEGHHKtH mechanism''~\cite{close2011infinity}
to fully credit other associated work.%
}
by adding four scalar fields with a potential that moves the ground state of
to a gauge-\emph{asymmetric} configuration.
Around this new ground state, linear combinations of the electroweak fields
acquire masses while ``eating up''~\cite{rubakov1999classical} three of the
four scalar fields to yield the three weak bosons and the photon:
$W^1$ and $W^2$ mix into $W^+$ and $W^-$, while $W^3$ and $B$ mix into $Z$ and
$\gamma$.
The fourth scalar is the Higgs boson $h$~\cite{
glashow1959renorm,
weinberg1967model,
salam1959weak,
rubakov1999classical,
cottingham2007greenwood
}.

Fermion masses arise through their Yukawa interactions with the Higgs field.
The subscript $_L$ on $SU\!(2)_L$ indicates that the weak interaction only
couples to fermions in left-chiral doublets: $(\nu_e, e)_L$,
$(u, d)_L$ and their repeats across the three generations.
In each generational trio (such as $u$, $c$, $t$), all three fermions share the
same quantum numbers, but have different masses.
A second generational difference occurs in quarks and neutrinos through the
difference between their weak flavour eigenstates and the mass eigenstates of
their free propagation;
flavour and mass eigenstates are linearly related through unitary matrices:
the CKM matrix for the quarks~\cite{
cabibbo1963unitary,
kabayasji1973cpv
}
and the PMNS matrix for the neutrinos~\cite{
maki1962remarks,
thomson2013modern
}.

Early versions of the Standard Model had only two fermion generations and
no neutrino masses~\cite{
wells2020discovery,
bjorken1985november
},
but the third generation and non-zero neutrino masses are now observed;
for example, the third generation is shown by $\mathrm{CP}$ asymmetry in
kaons~\cite{
cronin1964evidence,
kabayasji1973cpv
},
precise measurements of the $Z$ boson~\cite{
lep2006precision
},
and direct observations~\cite{
perl1977evidence,
herb1977observation,
abachi1995observation
};
neutrino masses are also necessary to explain their observed flavour
oscillations~\cite{
kamiokande1998measurement,
superk1998evidence,
superk1999measurement,
lsnd1998evidence
}.

Various theories can describe neutrino masses in the Standard Model, and it is
currently uncertain which form best describes nature;
this is both a limitation of the current Standard Model and an
opportunity for future refinement.
With future results, the Standard Model will continue to improve in its
correspondence with nature, and we discuss some such opportunities
for improvement in the next section.

\subsection{Opportunities for future Standard Models}
Problems with the Standard Model mean it needs replacing with supersymmetry
--- is a false statement.
Such an argument could be tempting to motivate this thesis, but the
Standard Model is actually very good within its domain.
Newton's laws are also very good within theirs \TODO{}

\TODO{Limitations: uncertainties and unknowns}
\TODO{uncertainties: lagrangian form for neutrino masses, mass values (hierarchy, mixing angles); composition of dark matter}
\TODO{unknowns: lack of convincing solutions; unification of gravity with quantum mechanics, origin of structure and parameters of standard model}

\paragraph{Neutrino masses.}
\TODO{Dirac neutrinos, possibly supplemented with a Majorana term}
\TODO{Weinberg operator for Majorana neutrinos}

\TODO{mass models: dirac, Majorana, seesaw~\cite{
wells2020discovery,
king2003neutrino,
thomson2013modern
}}

\paragraph{Gravitation and cosmology.}

\TODO{dark matter: rotation curves~\cite{
begeman1991rotation,
garrett2010dark
}}

\TODO{
primordial black holes~\cite{bernard2019primordial},
WIMPs~\cite{jungman1996wimp},
axions~\cite{kim2008axions},
dust?~\cite{garrett2010dark}
}
\TODO{combinations!}

\TODO{
dark matter,
cosmological constant~\cite{riess1998observational},
inflation~\cite{bezrukov2008higgs},
baryogenesis (matter-antimatter asymmetry)
}
\TODO{note \lcdm}

\paragraph{Origins.}

\paragraph{Aesthetics.}

\TODO{naturalness and fine-tuning~\cite{wells2020discovery}}
\TODO{strong CP problem}
\TODO{hierarchy problem}

\subsection{Supersymmetric models}
\label{sec:context_susy}

\TODO{higgs mass $\lesssim 130\,\eV[G]$~\cite{
quiros1997constraints,
wells2020discovery
}}

\TODO{primer~\cite{
martin2016primer
}}

% using https://www.tablesgenerator.com/latex_tables
\begin{table}[tp]
\centering
\begin{tabular}{rrc}
\multirow{5}{*}{bosons}   & \multirow{2}{*}{squarks}      & \makebox[5em]{\hfill $\tilde u$ \hfill $\tilde c$ \hfill $\tilde t$ \hfill}                  \\
                          &                               & \makebox[5em]{\hfill $\tilde d$ \hfill $\tilde s$ \hfill $\tilde b$ \hfill}                  \\[1ex]
                          & \multirow{2}{*}{sleptions}    & \makebox[5em]{\hfill $\tilde e$ \hfill $\tilde \mu$ \hfill $\tilde \tau$ \hfill}             \\
                          &                               & \makebox[5em]{\hfill $\tilde \nu_e$ \hfill $\tilde \nu_\mu$ \hfill $\tilde \nu_\tau$ \hfill} \\[1ex]
                          & Higgs                         & \makebox[7em]{\hfill ($h$) \hfill $H^0$ \hfill $A^0$ \hfill $H^\pm\!\!$ \hfill}             \\[3ex]
\multirow{3}{*}{fermions} & strong                        & \makebox[5em]{\hfill $\tilde g$ \hfill}                                        \\[1ex]
                          & \multirow{2}{*}{electroweak}  & \makebox[5em]{\hfill $\tilde C^\pm_1\!\!$ \hfill $\tilde C^\pm_2\!\!$ \hfill} \\
                          &                               & \makebox[7em]{\hfill $\tilde N_1\!$ \hfill $\tilde N_2\!$ \hfill$\tilde N_3\!$ \hfill$\tilde N_4\!$ \hfill} \\
\end{tabular}
\caption[Particles of the Minimal Supersymmetric Standard Model]{%
Particles of the Minimal Supersymmetric Standard Model
\TODO{relate to Standard Model}
\TODO{adds goldstino and gravitino}
\TODO{alternative $\chargino_i = \tilde C^\pm_i$, $\neutralino_i = N_i$}
\TODO{In this table, all bosons are spin-$0$ scalars and all fermions are spin-$1/2$.}
\TODO{The bracketed $h$ corresponds to the Standard Model Higgs boson.}
}
\label{tab:context_particles_mssm}
\end{table}

Table~\ref{tab:context_particles_mssm} \TODO{}

\subsubsection{Minimal supersymmetric models}
Lightest Supersymmetric Particle (LSP)

\subsubsection{Guage Mediated Supersymmetry Breaking}
Gauge Mediated Supersymmetry Breaking (GMSB)

\section{The \atlas\ Experiment}
\label{sec:context_atlas}
Towards explaining the main result of this thesis,
a search for effects from supersymmetric models
in data from the \atlas\ detector
on the Large Hadron Collider (LHC)~\cite{
atlas2022searches,
atlas2008experiment,
lhc2008machine
},
we first cover some relevant facts about the design and motivations
behind \atlas.

The \atlas\ detector was not delivered in a preordained form.
It was designed with intent to provide maximally useful and novel data within
real-world cost budgets.
Within general purpose utilities, \atlas's designs aimed to construct a
detector that could
\begin{itemize}
\item find the Higgs boson (which was predicted but yet unseen),
\item precisely measure known massive objects ($W$, $Z$, and $t$),
\item discover new (supersymmetric) particles~\cite{atlas1999design1}.
\end{itemize}
We now review these three goals with the subsequent news from \atlas's
operations.

\paragraph{Find the Higgs boson.}
Prior to \atlas's construction, the Higgs boson mass (and existence) was
poorly known, but loosely constrained by results like
$m(H) \gtrsim 90\,\eV[G]$ from searches at the
Large Electron–Positron Collider (LEP),
$m(H) = 76^{+85}_{-47}\,\eV[G]$ from other data on the electroweak sector,
and $m(H) \lesssim 1\,\eV[T]$ from theoretical unitarity arguments~\cite{
atlas1999design2,
ghinculov1998perturb,
lep1999ewk
}.

The Higgs boson couples to an extraordinary variety of fundamental particles.
For its discovery alone, \atlas's designers achieved sensitivity across the
entire $100$ to $1000\,\eV[G]$ window by targetting decays including
$H\to b\bar b$,
$H\to \gamma\gamma$,
$H\to ZZ^* \to 4\ell$, and
$H\to ZZ \to \ell\ell\nu\nu$;
these channels demand precise measurements of leptons, photons, and
$b$-jets (jets seeded from $b$ quarks), as well as sensitivity to the missing
transverse momentum of invisible neutrinos for $ZZ \to \ell\ell\nu\nu$ decays
at the very highest masses.

Both \atlas\ and \cms\ have now found a Higgs boson candidate at
$m(H) = 125\,\eV[G]$~\cite{
atlas2012higgs,
atlas2012combined,
cms2012higgs
}
and confirmed all of its tested Standard Model properties~\cite{
combined2016higgs,
atlas2022ten,
cms2022ten
}
such as its production mechanisms, decay rates,
spin, and parity~\cite{
HIGG-2013-01,
HIGG-2013-17,
HIGG-2014-06
}.

\paragraph{Precisely measure known massive objects.}
Other than the Higgs boson, the heavy Standard Model resonances are the
$W$ and $Z$ bosons and the top quark.
All of these are produced at high rates in LHC proton-proton collisions,
and present opportunities for \atlas\ to measure their properties
(such as masses and production cross-sections) with unprecedented precision.

Sensitivity to hadronic momenta is crucial,
both directly to measure heavy hadronic decays
(like $t \to bq\bar{q}$),
and indirectly for the missing transverse momenta from semi-invisible
leptonic decays (like $W \to \ell\nu$).
Since individual hadron identities are less relevant to these measurements,%
\footnote{%
Rare decays like
$W^\pm \to \pi^\pm \gamma$ or $W^\pm \to \pi^\pm \pi^+ \pi^-$
could give clean signals, but are not competitive due to their
tiny production rates and difficult identification~\cite{
cdf1996search,
mangano2014wpiy,
cms2021wpiy
}.
Leptonic $B$ hadron decays are useful in $m(t)$ measurements, but do not
require hadron identification~\cite{
CDF:2009mbf,
CMS:2016ixg,
ATLAS:2022jbw
}.%
}
it makes sense for \atlas's design to forfeit hadron identification systems
in favour of precise calorimetry with total angular coverage.

\atlas\ has now produced a plethora of competitively precise measurements
of the heavy Standard Model~\cite{atlas2021summarysm},
including $m(W)$~\cite{atlas2018wmass} and
$m(t)$~\cite{atlas2022symmarytop, atlas2019topmass, TOPQ-2015-03},
but also details of differential cross-sections~\cite{
STDM-2016-11,
STDM-2016-14,
TOPQ-2018-15,
TOPQ-2016-10
}.
Rare processes have also been studied in new levels of detail,
including
electroweak diboson ($WW/WZ/ZZ$ etc.)~\cite{
STDM-2015-21,
STDM-2015-23,
STDM-2017-09
}
and triboson~\cite{
STDM-2016-06,
STDM-2017-22,
STDM-2019-09
},
and rare top processes
($\ttbar Z$, $\ttbar W$ etc.)~\cite{
TOPQ-2013-05,
TOPQ-2018-01,
TOPQ-2020-03
},
These continue to not invalidate the Standard Model, and are aided in their
sensitivity by the lack of new phenomena,
which could have overpowered these rare signals.

\paragraph{Discover new (supersymmetric) particles.}
High-energy proton-proton collisions at the LHC therefore offer sensitivity
to new particles up to multi-TeV mass scales.
Heavy new particles are anticipated to either decay to known Standard Model
objects, or to be invisible in \atlas.
Designs that can discover such new particles are therefore neatly aligned with
our other goals of refining the current Standard Model
and to discovering the Higgs boson at any mass.
Visible decay products should already be caught by \atlas's abilities
to measure the Higgs boson and other Standard Model particles,
and the prospect of new invisible particles reinforces the need for precise
measurements of the missing transverse momentum.

Indeed, the main search presented in this thesis~\cite{atlas2022searches}
joins nicely with the pre-existing needs for sensitivity to signatures
including
$H \to b\bar b$
and
$ZZ \to \ell\ell\nu\nu$.

Dozens of searches for new phenomena have now been conducted in \atlas\ data,
with~\cite{
ATL-PHYS-PUB-2022-013
}
and without~\cite{
ATL-PHYS-PUB-2022-007,
ATL-PHYS-PUB-2022-012,
ATL-PHYS-PUB-2022-036,
ATL-PHYS-PUB-2022-034
} supersymmetry,
and show no convincing evidence for effects beyond the Standard Model.
This is not a failure. It is a success.
Our goal in science is not to discover something that is not there,
but to learn more about nature.
Finding no tangible new particles in LHC conditions is itself a discovery.

\subsection{Design}
Atlas is a Titan.
\atlas\ is the largest detector experiment at the LHC,
and alongside its biggest neighbours
\cms~\cite{cms2008experiment},
\lhcb~\cite{lhcb2008experiment},
and ALICE~\cite{alice2008experiment}
it is a multi-kilotonne beast,
and a wonder of scientific engineering~\cite{
atlas1994proposal,
atlas2008experiment,
atlas1999design1,
atlas1999design2
}.

Our search paper~\cite{atlas2022searches} contains concise and standard
descriptions of the \atlas\ detector and its relevant components.
Through its many inexact reproductions in past \atlas\ publications, the text
of these descriptions has become so highly refined that attempts to improve
it myself would probably be in vain.
Instead, this section quotes our published text with explanatory commentary.

\begin{figure}[tp]
\centering
\includegraphics[width=0.6\textwidth]{figures/atlas_cutaway_volume_1.pdf}
\caption[
A central slice of \atlas's design
]{%
A central slice of \atlas's design, reproduced from~\cite{atlas1999design1},
including a simulated event that interacts with several sensitive subsystems.
}
\label{fig:atlas_cutaway}
\end{figure}

\begin{quote}
``%
The ATLAS detector~\cite{atlas2008experiment} is a multipurpose particle
detector with a forward--backward symmetric cylindrical geometry and a near
$4\pi$ coverage in solid angle.%
\footnote{%
``ATLAS uses a right-handed coordinate system with its origin at the nominal
interaction point (IP) in the centre of the detector and the $z$-axis along the
beam pipe.
The $x$-axis points from the IP to the centre of the LHC ring, and the $y$-axis
points upwards.
Cylindrical coordinates $(r,\phi)$ are used in the transverse plane, $\phi$
being the azimuthal angle around the $z$-axis.
The pseudorapidity is defined in terms of the polar angle $\theta$ as
$\eta = -\ln \tan(\theta/2)$.
Angular distance is measured in units of
$\Delta R \equiv \sqrt{(\Delta\eta)^{2} + (\Delta\phi)^{2}}$.%
''\footnotemark~\cite{atlas2022searches}%
}
It consists of an inner tracking detector surrounded by a thin superconducting
solenoid providing a $2\,\mathrm{T}$ axial magnetic field, electromagnetic and
hadronic calorimeters, and a muon spectrometer.%
''~\cite{atlas2022searches}
\end{quote}
To illustrate this description, a vertical slice of \atlas\ is displayed in
Figure~\ref{fig:atlas_cutaway}, and each subsystem is described in more detail
in a subsection below:
the inner tracking detector in Section~\ref{sec:atlas_inner},
the electromagnetic and hadronic calorimeter systems in
Section~\ref{sec:atlas_calo},
and
the chambers of the muon spectrometer in Section~\ref{sec:atlas_muon}.
The final connection to the data analysed in this thesis is made through the
trigger system which is described in Section~\ref{sec:atlas_trigger},
and details of the data themselves are given in
Section~\ref{sec:atlas_data}

\footnotetext{%
This footnote appears in every \atlas\ publication, but was perfected
in~\cite{lester2013heffalon}.
True to form, internal \atlas\ guidelines ask that
``If you think you can improve on it, please consult with the
Publications Committee''~\cite{atlas_coordinate, atlas_footnote}.
}

\subsection{Inner detector}
\label{sec:atlas_inner}
The inner tracking detectors begin \atlas's measurements of new collisions
by precisely tracing the paths taken by energetic charged particles
as they curve through the solenoidal magnetic field.
Track curvature provides momentum information,
tracks' associations into collision vertices helps to separate the multiple
scatterings per crossing (pile-up),
and their associations into displaced vertices help to identify
decays.
\begin{quote}
``%
The inner tracking detector covers the pseudorapidity range $|\eta| < 2.5$.
It consists of silicon pixel, silicon microstrip, and transition radiation
tracking detectors.
An additional layer of silicon pixels, the insertable
B-layer~\cite{ATLAS-TDR-19, PIX-2018-001}, was installed before Run~2.%
''~\cite{atlas2022searches}
\end{quote}
This insertable B-layer is a silicon pixel detector that begins just
$3\,\textrm{cm}$ from the beam-pipe, and provides robust tracking for the
present and future high LHC luminosities,
particularly for sensitivity to particles whose lifetimes are just right to
decay inside the tracking system;
these `Goldilocks Zone' particles include B-hadrons, some tau leptons, and
maybe new species.

Following radially outwards, the silicon pixel and silicon microstrip%
\footnote{%
Formerly known as the SemiConductor Tracker~\cite{atlas1999design1},
the silicon microstrip system retains the acronym
SCT~\cite{atlas2008experiment}.%
}
detectors (which use silicon technology)
and the transition radiation tracking detectors (which uses `straw' tubes
gas tubes).

As for the whole \atlas\ detector, complete $\phi$ coverage and a large $\eta$
range minimizes leakage.
Towards \atlas's design goals, this inner tracking detector crucially provides
measurements of electrons, B-hadrons, and contributes to the reconstruction of
missing transverse momentum.

\subsection{Calorimeters}
\label{sec:atlas_calo}
Inelastic collisions cause energetic particles to scatter, distributing their
energy into macroscopic particle showers.
Calorimetry uses this principle to measure energies
by interleaving dense absorbing materials (which initiate showers)
with sensitive materials that convert showers into electrical signals.
In \atlas,
\begin{quote}
``%
Lead/liquid-argon (LAr) sampling calorimeters provide electromagnetic (EM)
energy measurements with high granularity.
A steel/scintillator-tile hadron calorimeter covers the central pseudorapidity
range ($|\eta| < 1.7$).
The end-cap and forward regions are instrumented with LAr calorimeters for both
the EM and hadronic energy measurements up to $|\eta| = 4.9$.%
''~\cite{atlas2022searches}
\end{quote}
The EM calorimeters provide located electron/photon energy measurements,
and the hadronic calorimeters subsequently aim to stop and measure hadronic
activity for later analysis in jet clustering.
Maximal angular coverage by the calorimeters further facilitates missing
transverse momentum estimates.

Neither `A' in \atlas\ stands for Argon, but perhaps one should.
Liquid-argon is the chosen active material in the EM and forward calorimeters
for its robust, linear response in converting showers to electrical signals
for readout by electrodes.
Liquid argon is poured as a liquid to fill the gaps between the various
metallic absorbers, which are folded into zigzag ``accordion'' shapes that
optimize uniformity around the detector.
The steel/scintillator hadronic tile calorimeter is a
``simple and cost effective''~\cite{atlas1996tile} design, which uses
photomultiplier tubes to convert light from its polystyrene scintillators into
electrical signals.
Particles lose some energy before reaching the calorimeters; a thin
liquid-argon pre-sampler is installed before the central EM calorimeter
to regain such loss.

These calorimeter systems achieve attenuations of at least ten interaction
lengths throughout their $|\eta| < 4.9$ range.
Maximizing attenuation minimizes noise from hadronic `punch-through' into the
muon chambers.

\subsection{Muon system}
\label{sec:atlas_muon}
Muons penetrate matter well because of their large mass and non-interaction
with the strong nuclear force.
\atlas\ doesn't try to stop energetic muons;
above around $3\,\eV[G]$, they pass right through the calorimeters, and are
instead measured by their curved tracks in a toroidal magnetic field that
envelopes dedicated detectors on \atlas's periphery.
\begin{quote}
``%
The muon spectrometer surrounds the calorimeters and is based on three large
superconducting air-core toroidal magnets with eight coils each.
The field integral of the toroids ranges between $2.0$ and
$6.0\,\mathrm{T\kern-0.15ex m}$ across most of the detector.
The muon spectrometer includes a system of precision chambers for tracking and
fast detectors for triggering.%
''~\cite{atlas2022searches}
\end{quote}
Muon subsystems come in four varieties, all of which operate by detecting
charges scattered by the muons as they fly through gaseous media.
Together, these systems track the muons' paths as the magnets bend them in the
longitudinal plane; these toroidal magnets generate fields outside the tile
calorimeter around the central barrel and in the end-caps as illustrated in
Figure~\ref{fig:atlas_magnets}.

\begin{figure}[tp]
\centering
\includegraphics[width=0.6\textwidth]{figures/atlas_magnets.pdf}
\caption[
Layout of \atlas's toroidal magnets around the tile calorimeter and inner
solenoid
]{%
Layout of \atlas's toroidal magnets around the tile calorimeter and inner
solenoid, including the eight barrel and end-cap coils,
reproduced from~\cite{atlas2008experiment}.
The \atlas\ muon spectrometer measures muons' tracks as they bend through
this peripheral magnetic field.
}
\label{fig:atlas_magnets}
\end{figure}

Measurements of the tracks' curvature yield estimates of the muons' momenta;
although extremely energetic muons will have straight tracks,
\atlas's construction and calibration achieves a high precision, which
increases to only $10\%$ relative error on tracks with $1\,\eV[T]$ transverse
momenta~\cite{
atlas1994proposal,
atlas2008experiment,
ATL-PHYS-PUB-2015-037
}.

\subsection{Trigger}
\label{sec:atlas_trigger}
The LHC's $25\,\mathrm{ns}$ time difference between proton bunches
(bunch spacing) is so fast that light-speed particles ejected from one crossing
are still zipping through \atlas's $12.5\,\textrm{m}$ radius when the next
collision arrives.
This extreme engineering is necessary to achieve the integrated luminosities we
need to see rare effects, but to avoid becoming overwhelmed
by data it also demands a strict data reduction mechanism.
This mechanism is the trigger:
\begin{quote}
``%
A two-level trigger system is used to select events.
The first-level trigger is implemented in hardware and uses a subset of the
detector information to accept events at a rate below $100\,\mathrm{kHz}$.
This is followed by a software-based trigger that reduces the accepted event
rate to $1\,\mathrm{kHz}$ on average depending on the data-taking conditions.%
''~\cite{atlas2022searches}
\end{quote}
The first-level (L1) trigger uses its fast preliminary information to accept
interesting event and mark their regions of interest.
Complete data in these regions of interest are passed to the second-level
(high-level trigger, HLT) to inform its decision of whether to accept each
event.
Events accepted by the HLT are recorded for future analysis~\cite{
atlas2016trigger,
atlas2008experiment
}.

\subsection{Data}
\label{sec:atlas_data}
Raw binary data are hard for us humans to interpret.
Before use in publishable analysis projects, software systems first mutate
those data into a workable form, of clean events with calibrated and
interpretable event variables.
\begin{quote}
``%
An extensive software suite~\cite{ATL-SOFT-PUB-2021-001} is used in the
reconstruction and analysis of real and simulated data, in detector operations,
and in the trigger and data acquisition systems of the experiment.%
''~\cite{atlas2022searches}
\end{quote}
Prediction is important in data analysis.
This software suite helps to generate predictive simulations, and provides
realism by passing both the real and simulated data through the same processing
pipelines~\cite{SOFT-2010-01, geant2003}.
Software models of the \atlas\ detector itself are refined with the results from
many calibration studies and in-situ measurements~\cite{atlas2008experiment},

The search presented in this thesis uses data from proton-proton collisions at
$\sqrt{s} = 13\,\eV[T]$ centre-of-mass energies in Run~2 of the
LHC, in which \atlas\ collected data in the years
$2015\textrm{--}2018$~\cite{atlas2022searches}.
This second run has various refinements over the Run~1, including beam energies
increasing from $\sqrt{s} = 7\textrm{ or }8\,\eV[T]$, and bunch spacings
halving from $50$ to $25\,\textrm{ns}$~\cite{lhc2006run2}.

Data collected under anomalous detector operations are hard to model.
Lists are maintained indicating which datasets arise from `good', normal
detector operations, and these lists are commonly employed to select only clean
subsets of good data for analysis projects.
Testament to the quality of \atlas's operations, this cleaning removes less
than $5\%$ of the Run~2 data~\cite{
DAPR-2018-01,
Golling:2011zy
}.

% closing comment?
Now that we have predictive models, in the Standard Model and supersymmetric
alternatives,
and interpretable data, from \atlas's measurements of LHC collisions,
we are ready to begin to squeeze some useful research from those data.

% this comment activates enlarged spacing for the final paragraph lol