### 2022-09-29 main/frontmatter.tex
first reasonable draft of acknowledgements

\titlepage[Homerton College]{%
A dissertation submitted to the University of Cambridge\\
for the degree of Doctor of Philosophy%
}

\begin{abstract}
Searching for new phenomena is an essential endeavour of experimental physics,
and is helped by guidance from plausible theoretical predictions.
This thesis presents such a search for signals of high-energy protons
scattering to massive resonances, whose dramatic decays are predicted by
supersymmetric models~\cite{atlas2022searches}.

Nothing new appears in the results, so they support the Standard Model in its
continuing successful extrapolation to describe the most extreme behaviours of
proton collisions.
Supersymmetry as a concept as not falsified by this lack of new phenomena,
nor by the many similar observations.
But it is also not confirmed.

Data are best interpreted with an understanding of their origins, so we begin
by introducing the theoretical and experimental background to our search, and
the subtleties of our data-analysis procedures in their historical context.
\end{abstract}

\begin{singlespacing}
\begin{declaration}
\noindent
\TODO{
This dissertation is the result of my own work, except where explicit
reference is made to the work of others, and has not been submitted
for another qualification to this or any other university.
This dissertation does not exceed the word limit for the respective
Degree Committee.
}
\vspace*{1cm}
\begin{flushright}
Rupert~Tombs
\end{flushright}
\end{declaration}
\end{singlespacing}

\begin{acknowledgements}
Warning: many thank-yous incoming.
Friends and family made these years worthwhile;
deepest love and thanks to all of you who have come and gone.

\paragraph{958 2018/19.}
Thank you Daniel~Noel for the years of comradery and surviving this with me
--- I hope you enjoyed crushing me OTB.
Thank you Holly~Pacey, not only for non-sofa antics
--- you are an inspirational human and academic and you deserve endless thanks
for putting up with the ravings of my most miserable months on this
project\ldots\ oops!

\paragraph{Supervision.}
Thank you Christopher~Gorham~Lester for guidance and encouragement,
and for frequent reminders of the fun and intrigue outside \atlas.
Thank you Sarah~Williams for your advice and support inside \atlas, and
for the invaluable experience of teaching at Queens' College.

\paragraph{\atlas.}
Thank you to the Cambridge \atlas\ group and especially Tina~Potter for
sharing help and wisdom.
Thanks to all collaborators mentioned in Chapter~\ref{chapter:2ljets},
to Ben~Hooberman and Sarah~Williams for leadership, and especially to
Jon~Long for our teamwork in the push for publication,

\paragraph{Misc.}
Thank you Christopher~Gorham~Lester, Holly~Pacey, and Daniel~No\"el for your
stunningly precise and valuable comments on drafts of this thesis.
Thank you to Homerton College and Melanie~Keene.
This document uses parts of `hepthesis'~\cite{hepthesis}.
\end{acknowledgements}

\begin{preface}
\begin{singlespacing}
\begin{epigraphs}
\qitem{%
Begin reading at Chapter N. Do \emph{not} read the quotations
that appear at the beginning of the chapter.%
}%
{Donald~E.~Knuth,
\textit{The Art of Computer Programming, Volume~1},
1975~\cite{knuth1975art}}
\end{epigraphs}
\end{singlespacing}

Data Intensive Science in High-Energy Physics

\TODO{Context of Standard Model theory, directions in which future developments
could improve it, and supersymmetric models}
supersymmery (SUSY)
Chapter~\ref{chapter:theory}

\TODO{the design of the \atlas\ detector}
Chapter~\ref{chapter:experiment}

\TODO{theory and practice of data science, as it could be for clarity
and how it is in our \atlas\ SUSY conventions}
Chapter~\ref{chapter:searches}

\TODO{the main event: the search with many of its gory details in addition
to those presented publically in the paper~\cite{atlas2022searches}}
Chapter~\ref{chapter:2ljets}

\paragraph{Other work.}
\TODO{Use some of the data analysis ideas of Chapter~\ref{chapter:searches}
to build predictive models
and test them to test for symmetries}
\TODO{citations}

\end{preface}

\afterpage{%
\phantomsection
\addcontentsline{toc}{chapter}{Contents}%
}
\tableofcontents

% the following blank page should also have no page number
\thispagestyle{empty}